%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beginnin stuff
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\author{Florian Kaufmann}
\title{Bash Summary}
\date{\today}
\maketitle
\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

QtCurve a) is a set of widget styles for Qt4/KDE4 / KDE4 / Gtk2 based
Apps b) is a desktop theme for GTK+ and Qt widget toolkits

A (Desktop) Theme is preset set of graphical options. So all options as
  a whole are affected, which might be also a disadvantage. Gnome/KDE
  have 2 Themes.

A widget toolkit (=widget library, GUI toolkit): A set of widgets. The
provided widgets typically adhere to a unified design specification.
Examples: Qt, GTK+, Motif, Xaw, MFC, AWT, Swing.

Window manager
  Compositing: Metacity, Quartz (Mac OS X), Xfwm, KWin, Compiz
  Stacking window managers: 
  Tiling:

gtk+ (=the gimp toolkit) is cross platform widget toolkit
gtk+2 superceeds gtk+1. 

  
Metacity: Compositing window manager, included in GNOME.
Xfwm: Compositing window manager, Xfce's window manager.
Quartz Compositor: Composing window manager of Mac OS X
Desktop Window Manager (=DWM): Compositing window manager, part of
Windows Vista.

Window decorations are provided by the window manager. They typically
consist of a title bar, buttons for closing/mini|maximizing, and a small
border around the other three sides.

Xlib: Interact with X11 server
  Xt(=Xi,X toolkit,Intrinsics): Create and use widgets
    Xaw (a widget toolkit): Specific widgets
    Motif (a widget toolkit): Dito
  GTK+ (a widget toolkit)
  Qt (a widget toolkit)
    

Widget(=Control): Element of a GUI providing a) a single interaction point
for the direct manipulation of a given kind of data b) a container for
other widgets.

\end{document}
%%%
